---
title: "Statistical Methods"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Statistical Methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

The IKDDS dashboard uses several statistical methods for centre
comparison. This vignette demonstrates each method with worked
examples using synthetic data.

```{r}
library(ikdds.dashboard)

df <- generate_synthetic_data(seed = 42, n_total = 1200)
```

## Wilson Score Confidence Intervals

For proportion-based metrics (e.g., "% of patients with URR >65%"),
the dashboard uses Wilson score confidence intervals rather than the
normal approximation. This provides better coverage for small samples
and avoids intervals outside [0, 1].

### The `wilson_ci()` Function

```{r}
# 15 out of 20 patients achieving target
wilson_ci(x = 15, n = 20)
```

```{r}
# Compare small vs large sample
wilson_ci(x = 3, n = 5)    # Small sample — wide interval
wilson_ci(x = 60, n = 100) # Large sample — narrow interval
```

### Applied to Centre Data

```{r}
# Compute URR achievement with CIs
urr <- compute_urr_achievement(df)
urr[, c("centre_code", "n", "x", "proportion", "lower_ci", "upper_ci")]
```

## Caterpillar Plots

Caterpillar plots display centre-level results as ordered
dot-and-whisker plots, with a national average reference line:

```{r fig.height=4}
urr <- compute_urr_achievement(df)
plot_caterpillar_proportion(urr, target = 0.65)
```

### Median Caterpillar

For continuous metrics like URR median:

```{r fig.height=4}
urr_med <- compute_urr_median(df)
plot_caterpillar_median(urr_med)
```

## Spiegelhalter Funnel Plots

Funnel plots show each centre plotted as sample size (x-axis) vs
achievement rate (y-axis), with control limits based on expected
binomial variation around the national mean.

### The `funnel_limits()` Function

```{r}
# Compute funnel limits for a range of sample sizes
limits <- funnel_limits(
  target_rate = 0.70,  # National proportion
  n_range = c(20, 300)
)
head(limits)
```

### Applied to Metrics

```{r fig.height=4}
urr <- compute_urr_achievement(df)
plot_funnel(urr)
```

Centres outside the 95% limits (solid lines) show more variation
than expected by chance. Centres outside the 99.7% limits (dashed)
are considered statistical outliers.

## Suppression Engine

The dashboard suppresses results for centres with fewer than 10
patients to prevent identification risk:

```{r}
# Create a small test dataset
small_df <- df[df$centre_code %in% c("BEA", "MAT"), ]

# The engine pipeline handles suppression
urr <- compute_urr_achievement(small_df)

# Apply suppression (centres with n < 10 get NA values)
suppressed <- apply_suppression(urr, min_n = 10)
suppressed[, c("centre_code", "n", "proportion", "suppressed")]
```

## Completeness Labelling

Centre labels are annotated with their data completeness rate.
If completeness falls below 70%, a warning is added:

```{r}
urr <- compute_urr_achievement(df)
labelled <- add_completeness_labels(urr, df, "qblg9")
labelled[, c("centre_code", "n", "label")]
```

## Median Confidence Intervals

For continuous metrics, the dashboard uses order-statistic CIs
for the median via `median_ci()`:

```{r}
# Median CI for a sample
set.seed(42)
x <- rnorm(50, mean = 70, sd = 5)
median_ci(x)
```

This uses the exact binomial method to find the order statistics
that bound the 95% CI for the population median.

## National Averages

National averages are computed as weighted averages across centres:

```{r}
urr <- compute_urr_achievement(df)

# Proportion: total events / total patients
national_average_proportion(urr)

# Median: sample-size-weighted mean of centre medians
urr_med <- compute_urr_median(df)
national_average_median(urr_med)
```

## Individual Patient QA Medians

The Individual Patient QA page uses simpler inline computations
rather than the metrics engine. For each of 11 continuous metrics,
it computes:

- **Patient value** — the individual patient's recorded result
- **Centre median** — `median(x, na.rm = TRUE)` across all patients
  at the same centre
- **National median** — `median(x, na.rm = TRUE)` across all patients
  nationally

This direct approach avoids the suppression/completeness pipeline
since the purpose is clinical QA rather than centre benchmarking.

```{r}
# Example: compute medians for URR
centre_urr <- df$qblg9[df$centre_code == "BEA" & !is.na(df$qblg9)]
national_urr <- df$qblg9[!is.na(df$qblg9)]

cat("Centre median URR (BEA):", median(centre_urr), "\n")
cat("National median URR:    ", median(national_urr), "\n")
```
