---
title: "Architecture"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Architecture}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This article is for people changing the package, rather than using it. It
describes how a number gets from the data source onto a plot, and where to
intervene when you need to change one.

The short version: the app is **data-driven from two registries plus a compute
engine**. Almost nothing about a metric is hard-coded in the Shiny layer —
adding one is a YAML entry plus a function, and the engine does the rest.

## The four layers

### 1. Data source

`R/data-source.R`, `R/data-synthetic.R`, `R/data-redcap.R`.

`get_audit_data(config)` is a factory returning a tibble with one row per
patient. Both the synthetic and the REDCap source produce the *identical*
column schema, defined by `audit_data_columns()` and enforced by
`validate_audit_data()`. That equivalence is what lets every downstream layer
be tested without credentials.

Column names are raw REDCap field codes — `qblg9` (URR), `qble1` (Hb), `qhd20`
(vascular access). See the `get_audit_data()` roxygen for the full mapping.

The codes look unreadable and stay anyway: they are the field names in the UKRR
v5.1 REDCap project, so a metric definition can be checked against the canonical
dictionary in `ikdds-etl` by eye. Renaming them here would break that link and
hide drift rather than fix it.

### 2. Metrics registry

`inst/metrics/metrics.yml` → `R/metrics-registry.R`.

The YAML is the single source of truth for each metric's title, domain, target,
`value_field`, suppression threshold, and — critically — `compute_fn`, the
*name* of the R function that computes it.

`metrics_registry()` reads and caches the YAML. Call `reset_registry_cache()`
after editing it in a live session, or you will be debugging a stale metric that
the file on disk no longer describes.

`value_field` codes are not ours to invent: they must exist in the canonical
dictionary held by `ikdds-etl`. After editing the YAML, check it:

```bash
Rscript ../ikdds-etl/tools/check_dictionary_drift.R inst/metrics/metrics.yml
```

The older NESKDDS sample REDCap project and the CRF/DRS drafts use a different
coding scheme with zero overlap with this one. If you find a code there, it is
historical record, not a spec — see `ikdds-etl` and its `DRIFT_REPORT.md`.

### 3. Metrics engine

`R/metrics-engine.R`.

`compute_with_engine(data, metric_id, group_col)` is the standard pipeline:

1. look up the metric config in the registry;
2. resolve `compute_fn` by name from the package namespace and call it;
3. add completeness labels;
4. apply small-n suppression.

`group_col` is `"centre_code"`, `"consultant"`, or `"region"` — the "compare by"
toggle in the UI.

**Suppression is a disclosure control, not a plotting nicety.**
`apply_suppression()` (default `min_n = 10`, set per metric in the YAML) sets
the proportion, median, CIs *and `n` itself* to `NA` for under-sized groups.
`n` is included deliberately: there are 23 centres and several are small
satellite units, so "2 of 3 patients" is potentially patient-identifying even
with no identifier attached to it.

Going around the engine is how suppression gets lost. A compute function that a
page calls directly is a bug even when the numbers look right, and so is
recomputing a count from the raw tibble for a label or a tooltip.

### 4. Domain compute functions

`R/metrics-{domain}.R` — one file per clinical domain (dialysis, bp,
biochemistry, anaemia, access, demographics, prd, data quality).

These return tidy tibbles via the shared builders in `R/metrics-helpers.R`:
`compute_centre_proportion()` and `compute_centre_median()`, which carry Wilson
CIs, median CIs, and funnel limits. A function named by `compute_fn` must accept
`data` and optionally `group_col`.

Use the shared builders rather than a bespoke `mean()`. Wilson intervals are
there because centre denominators are small and a normal-approximation interval
misbehaves exactly where this audit is most sensitive — a small unit near 0% or
100%. An interval that is wrong in that corner is an outlier accusation against
a named hospital. `vignettes/statistical-methods.Rmd` covers the statistics in
full.

## Missingness is a finding, not a nuisance

`add_completeness_labels()` puts "(X% missing)" on the centre label on purpose.
A centre that does not report URR and a centre with poor URR must not look the
same on a plot.

Do not drop incomplete rows to make a chart tidier, and do not impute anything:
the audit reports what was recorded. `vignettes/data-quality-workflow.Rmd`
describes how completeness is surfaced to the data manager.

## Adding or changing a metric

1. Add or edit the entry in `inst/metrics/metrics.yml`.
2. Write the matching `compute_*` function in the relevant
   `R/metrics-{domain}.R` — or reuse an existing one — and `@export` it so the
   engine can resolve it by name.
3. If the metric appears on the spec/handover, register it in
   `dashboard_registry()` too (see below).
4. Run the drift check, then `devtools::document()` and `devtools::test()`.

```{r}
# After editing the YAML in a live session
reset_registry_cache()
compute_with_engine(get_audit_data(), "urr_achievement", "centre_code")
```

## The Shiny layer

`R/ui.R` and `R/server.R` wire everything together. Each page is a Shiny module
pair, `mod_{domain}_ui()` / `mod_{domain}_server()`, in `R/mod-{domain}.R`.

`mod_filters` owns the cascading centre/consultant filters and exposes the
reactives `filtered_data`, `compare_by`, and `active_filters`, which
`app_server()` passes down to every module.

**Role-based row filtering happens once**, in `app_server()` via
`filter_by_role()` (`R/access-control.R`), before data reaches any module. Do
not re-filter inside a module, and do not pass unfiltered data past that point:
a clinical user must not be able to reach another centre's rows through a module
that forgot.

## Spec / Power BI handover

`R/spec-registry.R` (`dashboard_registry()`) is a *separate* declarative
registry describing every view, card, plot, variable mapping (REDCap ↔ eMed/IIS
columns), and pseudocode. `R/spec-generate.R` turns it into a multi-sheet Excel
workbook that lets the HSE IIS team recreate the dashboard in Power BI.

It is independent of `metrics.yml` and must be kept in sync **manually** when
views change. It is the document another team builds from, so a view changed
here and not there becomes their bug, months later, with no way for them to know
it drifted. If you change a view and cannot update the registry, say so
explicitly rather than leaving it.
